---
title: "Overview of Automated Table 1 Generation"
author: "Ruyi Pan, Siyue Yang, Jessica Gronsbell"
date: "20/07/2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r include=FALSE}
# Load table 1 functions.
source("generate_table1.R")

# Library to knit html tables.
library(knitr)
library(kableExtra)
```

This document uses a simulated data set to illustrate the functionality of the table 1 function.

```{r}
set.seed(92047)

# Simulate example data.
example_data <- simulate_table1_data()

# Quick summary of the data.
str(example_data)
```

### Default Table 1 with the example data set. 

```{r}
# The default table 1 summarizes across all the variables in the dataframe.
# The displayed variable names correspond to the column names in the dataframe.
tab1_basic <- generate_table1(patient_level_data = example_data)

tab1_basic
```

```{r}
tab1_basic %>% save_kable(file = "tab1_basic.png")
```


### Functionality 1: Change the number of significant digits.

```{r}
# The default number of significant digits for numeric value (as in example above) is 3.
# The default number of significant digits for proportion (as in example above) is 2.
tab1_sigdig <- generate_table1(
  patient_level_data = example_data,
  num_digits = 2,
  pct_digits = 2
)

tab1_sigdig
```

```{r}
tab1_sigdig %>% save_kable(file = "tab1_sigdig.png")
```


### Functionality 2: Add a stratifying variable. 

```{r}
tab1_strat <- generate_table1(
  patient_level_data = example_data,
  stratifying_var = "Variable_6"
)

tab1_strat
```

```{r}
tab1_strat %>% save_kable(file = "tab1_strat.png")
```

### Functionality 3: Rename variables.

```{r}
# Create a list to rename desired variables.
example_dictionary <- list(
  Variable_1 = list(rename = "Z-score"),
  Variable_3 = list(rename = "Smoking Status"),
  Variable_6 = list(rename = "Neighborhood")
)

tab1_rename <- generate_table1(
  patient_level_data = example_data,
  var_dictionary = example_dictionary,
  stratifying_var = "Variable_6"
)

tab1_rename
# Note: You can also rename variables by renaming the column in the dataframe.
```

```{r}
tab1_rename %>% save_kable(file = "tab1_rename.png")
```

### Functionality 4: Specify variable type.

```{r}
# Change type for Variable_5.
# The default sets a variable as categorical if the number of unique values is
# less than log(n) where n is the number of rows in the dataframe.
example_dictionary <- list(
  Variable_1 = list(rename = "Z-score"),
  Variable_3 = list(rename = "Smoking Status", type = "categorical"),
  Variable_2 = list(rename = "Age", type = "categorical")
)

tab1_type <- generate_table1(
  patient_level_data = example_data,
  var_dictionary = example_dictionary,
  stratifying_var = "Variable_6"
)

tab1_type
```

```{r}
tab1_type %>% save_kable(file = "tab1_type.png")
```

### Functionality 5: Specify variable units.

```{r}
# Create a list to rename desired variables, adding units for height.
example_dictionary <- list(
  Variable_1 = list(rename = "Z-score"),
  Variable_3 = list(rename = "Smoking Status"),
  Variable_5 = list(rename = "Height", unit = "cm")
)

tab1_units <- generate_table1(
  patient_level_data = example_data,
  var_dictionary = example_dictionary,
  stratifying_var = "Variable_6"
)

tab1_units
```

```{r}
tab1_units %>% save_kable(file = "tab1_units.png")
```

### Functionality 6: Specify table style.

```{r}
example_dictionary <- list(
  Variable_1 = list(rename = "Z-score"),
  Variable_3 = list(rename = "Smoking Status")
)

tab1_style <- generate_table1(
  patient_level_data = example_data,
  var_dictionary = example_dictionary,
  stratifying_var = "Variable_6",
  my_table_style = "Rtable1-zebra"
)
```

```{r}
tab1_style %>% save_kable(file = "tab1_style.png")
```

### Example: Using the above functionality with an existing data dictionary.

```{r}
my_full_dictionary <- read.csv("Example_Data_Dictionary.csv",
  row.names = 1,
  na.strings = c("", "NA")
)

# Subset the dictionary to variables in the given data set.
my_dictionary <- my_full_dictionary[colnames(example_data), ]

# Transform the dictionary into a list for use in the table 1 function.
colnames(my_dictionary) <- c("rename", "type", "unit")

# Function change dataframe dictionary to list dictionary.
my_dictionary_list <- data_dictionary_2_list(my_dictionary)

generate_table1(example_data, var_dictionary = my_dictionary_list)
```

### Additional directions.
To get your table into word, knit a markdown file to HTML and copy the outputted table to the word document. 
